%! Author = dario
%! Date = 7/27/24

% Preamble
\documentclass[11pt]{article}

\title{Optimal Transport}
\author{Dario Filatrella}
\date{July 2024}
% Packages
\usepackage{amsmath}

% Document
\begin{document}
    \maketitle


    \section{Equiboundedness of length of fibers}

    What we want to show is that, for fixed $\alpha$, $\mu^+$ and $\mu^-$, the optimal transport parametrization $\chi$
    is equibounded in the sense that there is no path $\gamma$ that is non self-intersecting and such that $|\gamma(t)|_\chi$ is decreasing and positive
    for every $t$.
    The idea of the proof is to assume that there is such a path and show that it is not optimal by finding a better competitor.
    Without loss of generality, we can assume that $\gamma$ is parametrized by arc length and that $\gamma(0) = 1$.
    We define a competitor $\tilde\chi$ as the union of two paths, $\tilde{\gamma}$ and $\tilde{\psi}$, defined as follows:
    \begin{align*}
        |\tilde{\gamma}(t)| &= |\gamma(t)| - |\gamma(K)| \quad \text{for } t \in [0, K] \\
        |\tilde{\gamma}(t)| &= |\gamma(t)| \quad \text{for } t \in [K, \infty) \\
        \tilde{\gamma}(t) &= \gamma(t) \quad \text{for } t \in [0, \infty) \\
        |\tilde{\psi}(t)| &= \gamma(K) \\
        \tilde{\psi}(t) &= \text{Straight line from } \gamma(0) \text{ to } \gamma(K)
    \end{align*}

    Basically, we cut from the beginning of $\gamma$ the part that is longer than $|\gamma(K)|$ and we add a straight line from
    $\gamma(0)$ to $\gamma(K)$. We can arbitrarily start from 0 as we can just shift the path to the right and look at the restriction,
    rescaling it so that it has mass 1.

    We can now compute the cost of $\tilde{\chi}$:

    \begin{align*}
        E^\alpha(\tilde{\chi}) - E^\alpha(\chi) &= \int_0^K |\psi(t)|^\alpha \psi'(t) \, dt - \int_0^K \left[\gamma(t) - \gamma(K)\right]^\alpha \, dt \\
        &\quad - \int_0^K \gamma(t)^\alpha \, dt \\
        &\le |\gamma(K)|^\alpha diam(X) +  \int_0^K \left[\gamma(t) - \gamma(K)\right]^\alpha \, dt \\
        &\quad - \int_0^K \gamma(t)^\alpha \, dt \\
    \end{align*}

    We can now use the concavity of the function $x \mapsto x^\alpha$ to expand to the first order the second term in the right hand side:

    \begin{equation}
        \int_0^K \left[\gamma(t) - \gamma(K)\right]^\alpha \, dt \le \int_0^K \gamma(t)^\alpha - \alpha \gamma(t)^{\alpha - 1} \gamma(K) \, dt
    \end{equation}

    by linearity of the integral, two terms cancel out and we are left with:

    \begin{equation}
        E^\alpha(\tilde{\chi}) - E^\alpha(\chi) \le |\gamma(K)|^\alpha diam(X) - \alpha \int_0^K \gamma(t)^{\alpha - 1} \gamma(K) \, dt
    \end{equation}

    We are left to prove that this is negative for some $K$. Here I got stuck, if we assume that there exist $\epsilon,\epsilon' > 0$ such that
    $$t^{-\frac{1}{\alpha} - \epsilon'} \le |\gamma(t)| \le t^{-\frac{1}{\alpha} + \epsilon}$$
    For convenience, we assume that the path starts from 1 instead of 0.
    then the following computations will yield the desired result. However, I am not sure how to proceed for functions that decrease faster or slower than any power law.
    I have tried expanding to first and second order the integral at different points but without any good result. What often happens is that
    after some bounding, I get to an inequality that is too strong and fails for exponentials.

    Here there are the computations for the power law case (setting $d=diam(X)$):
    \begin{align*}
        E^\alpha(\tilde{\chi}) - E^\alpha(\chi) &\le d|\gamma(K)|^\alpha - \alpha \int_0^K \gamma(t)^{\alpha - 1} \gamma(K) \, dt \\
        &\le dK^{(-\frac{1}{\alpha} - \epsilon)\alpha} - \alpha\gamma(K)\int_1^K t^{(-\frac{1}{\alpha} - \epsilon)(\alpha - 1)} \, dt \\
        &\le dK^{-1 - \epsilon\alpha} - \frac{\alpha}{-\epsilon\alpha + 1/\alpha + \epsilon} \left( \gamma(K) K^{-\epsilon\alpha + 1/\alpha + \epsilon} - \gamma(K) \right) \\
        &\le dK^{-1 - \epsilon\alpha} - \frac{\alpha}{-\epsilon\alpha + 1/\alpha + \epsilon} \left( K^{-\epsilon\alpha + \epsilon - \epsilon'} - K^{-1/\alpha + \epsilon} \right) \\
    \end{align*}

    Observe that $\epsilon$ and $\epsilon'$ can be arbitrarily close which implies that $-\epsilon\alpha + \epsilon - \epsilon' > -1/\alpha - \epsilon$.
    This means that for $K$ sufficiently large the sum of the two terms will be negative and we will have the contradiction.

    We did not really use that the domain $X$ is compact, apart from the fact that the diameter is finite. I tried to use the fact
    that the path $\gamma$ will get arbitrarily close to itself but without much success. I got to this result: //
    Let $\gamma$: [0, $\infty$) $\to$ X be a continuous path. Then there exists $x \in X$ and $(t_n)_{n \in  \mathcal{N}} \to \infty$ such that
    $\gamma(t_n) \to x$.
    Proof: Consider the set $\{\gamma(n): n \in \mathcal{N}\}$, by compactness we can extract a subsequence that converges to some $x$.

    This also allows us to do the "cut and jump" around a fixed point, but it didn't seem to help. Another thing I tried is to quantify how often
    it can pass again but I did not find any known result online, and by a volume argument it seems that the time between intersections is too large.
    Obviously, this cut would not be optimal either as if we zoom close enough this path should look like a Lebesgue segment
    and its optimal transport parametrization is more complex than a straight line but I wouldn't know how to use this.

\end{document}